---
title: "Initial Results"
format: html
editor: visual
---

Loading up the data

```{r}
library(haven)
library(readxl)

gsp_public  <- read_dta("gsp_public_baseline_PaperVersion.dta")
school_gps <- read_xls("school gps.xls")


```

Now Cleaning it, calculating distance and merging

```{r}
library(tidyverse)
library(dplyr)
library(geosphere)

school_gps <- school_gps |> mutate(
  north = as.numeric(north), 
  n_degree = as.numeric(n_degree)
  )

GPS_data <- mutate(school_gps, 
lat = n_degree + north/60, 
lon = e_degree + east/60)            

ref <- GPS_data |> filter(schoolid == 916)

dist_out <- GPS_data %>%
  mutate(
    dist_lig = distHaversine(
      cbind(lon, lat),
      c(ref$lon[1], ref$lat[1])
    ) / 1000
  ) %>%
  arrange(dist_lig)

main <- gsp_public |> rename(school_id = school_baseline)
dist_out <- dist_out |> rename(school_id = schoolid)

main <- merge(main, dist_out, by = "school_id")


                   
```

Creating Table 1

```{r}
df <- main %>% filter(restricted_sample == 1)

# variables to summarize (must be numeric or 0/1)
vars <- c("score_2000", "score_2001", "score_2002", "missing_score_2001", "dist_lig", "bs_questionnaire_result", "treatment", "teso", "ateso_percent")

summ_one <- function(x, name) {
  tibble(
    Variable = name,
    N        = sum(!is.na(x)),
    Missing  = sum(is.na(x)),
    Mean     = mean(x, na.rm = TRUE),
    SD       = sd(x, na.rm = TRUE),
    Min      = min(x, na.rm = TRUE),
    Median   = median(x, na.rm = TRUE),
    Max      = max(x, na.rm = TRUE)
  )}


table1 <- map_dfr(vars, ~ summ_one(df[[.x]], .x)) %>%
  mutate(across(c(Mean, SD, Min, Median, Max), ~ round(.x, 3)))

table1
```

Now to upload this table into overleaf

```{r}
library(knitr)

tab_tex <- knitr::kable(
  table1,
  format = "latex",
  booktabs = TRUE,
  caption = "Summary statistics",
  label   = "tab:summary",
  align   = c("l", rep("r", ncol(table1)-1)),
  col.names = c("Variable","N","Missing","Mean","SD","Min","Median","Max")
)
cat(tab_tex, file = "table1.tex")
```

This is to push the table 1 into overleaf

```{bash}
cat > .gitignore << 'EOF'
.ipynb_checkpoints/
.Rhistory
.Rproj.user/
__pycache__/
*.log
*.aux *.bbl *.blg *.out *.toc
*.dta
*.xls
outputs/*.tmp
data_raw/
EOF

mkdir -p outputs
mv table1.tex outputs/table1.tex

git add InitialResults.qmd outputs/table1.tex .gitignore
git status

git commit -m "Add InitialResults.qmd and outputs/table1.tex; add .gitignore"

git pull --rebase origin main

git push origin main
```

Now a graph demonstrating the main correlation in data

```{r}
library(ggplot2)

main_teso <- main |> filter(teso == 1, restricted_sample == 1) 

p <- ggplot(main_teso, aes(x = dist_lig, y = score_2002)) +
  geom_point(aes(color = factor(treatment, levels = c(0,1),
                                labels = c("Control","Treatment"))),
             alpha = 0.6) +
  geom_smooth(aes(color = factor(treatment, levels = c(0,1),
                                 labels = c("Control","Treatment")),
                  linetype = factor(treatment, levels = c(0,1),
                                    labels = c("Control","Treatment"))),
              method = "lm", se = FALSE, linewidth = 1) +
  labs(
    x = "Distance to lightning (km)",
    y = "Test score (2002, normalized)",
    color = "Group",
    linetype = "Group",
    title = "Distance vs. 2002 Test Score by Treatment"
  ) +
  theme_minimal()

# Preview
p
```

Pushing it to Overleaf

```{r}
if (!dir.exists("outputs")) dir.create("outputs")
ggsave("outputs/fig_distlig_score2002_teso1.pdf", p, width = 6.5, height = 4.2)
ggsave("outputs/fig_distlig_score2002_teso1.png", p, width = 6.5, height = 4.2, dpi = 300)
```

```{bash}
git add outputs/fig_distlig_score2002_teso1.*
git commit -m "Add dist_lig vs score_2002 plot (teso==1, restricted sample) by treatment"
git pull --rebase origin main
git push origin main
```

Now to the OLS regression

```{r}
install.packages("modelsummary")
library(modelsummary)
dir.create("Proposal-2/outputs", showWarnings = FALSE, recursive = TRUE)

lm1 <- lm(main_teso, formula = score_2002 ~ dist_lig*treatment)

modelsummary(
  lm1,
  coef_map = c("(Intercept)"="Constant","treatment"="Treatment",
               "dist_lig"="Distance to lightning (km)",
               "treatment:dist_lig"="Treatment × Distance"),
  statistic = "({std.error}) [{p.value}]",
  gof_map = c("nobs","r.squared","adj.r.squared"),
  title = "OLS (i.i.d. standard errors)",
  notes = "Standard errors assume i.i.d. errors.",
  output = "Proposal-2/outputs/table_reg.tex")

```

```{r}
library(modelsummary)

lm1 <- lm(score_2002 ~ dist_lig * treatment, data = main_teso)
options(modelsummary_format_numeric_latex = "siunitx")
mdv <- mean(main_teso$score_2002, na.rm = TRUE)

tex <- modelsummary(
  list("(1)" = lm1),
  coef_map = c(
    "(Intercept)"        = "Constant",
    "dist_lig"           = "Distance to lightning (km)",
    "treatment"          = "Treatment",
    "dist_lig:treatment" = "Treatment × Distance"
  ),
  estimate  = "{estimate}{stars}",
  statistic = "({std.error})",
  stars     = c("*"=.10, "**"=.05, "***"=.01),
  fmt       = 3,
  gof_map   = c("nobs"="Sample size", "r.squared"="$R^2$", "adj.r.squared"="$R^2_{adj}$"),
  add_rows  = data.frame(term = "Mean of dependent variable", `(1)` = sprintf("%.3f", mdv)),
  output    = "latex"   # returns an S4 object
)

cat(
  "\\begin{table}[!htbp]\\centering\\small\\begin{threeparttable}\n",
  "\\caption{OLS: score$_{2002}$ on Treatment, Distance, and Their Interaction}\n",
  "\\label{tab:score2002_treat_dist}\n",
  as.character(tex),   # <<-- coerce to character
  "\n\\begin{tablenotes}[flushleft]\\footnotesize\n",
  "\\item Notes: i.i.d. standard errors in parentheses. Stars: * $p<0.10$, ** $p<0.05$, *** $p<0.01$.\n",
  "\\end{tablenotes}\n\\end{threeparttable}\\end{table}\n",
  file = "outputs/table_reg_journal.tex", sep = ""
)

```

Now I need to pass it to overleaf

```{bash}
git add outputs/table_reg_journal.tex
git pull --rebase origin main     # bring remote changes on top of yours
# if conflicts: fix files, then -> git add <fixed-files>; git rebase --continue
git push origin main
```

Extra

Now this is a test for a possible endogenous variable in an IV analysis, just with distance and questionnaire dist_lig was significant, however when controlling for score_2000 it lost its predictive power, which makes me believe that its had a lot of OVB.

```{r}

main_teso <- main |> filter(
  teso == 1, restricted_sample ==1)

m1 <- lm(formula = bs_questionnaire_result ~ dist_lig*treatment + score_2000, data = main_teso)
m2 <- lm(formula = score_2002 ~ bs_questionnaire_result + score_2000, data = main_teso)
m2001 <- lm(score_2001 ~ dist_lig*treatment, data = main_teso)
m2002 <- lm(score_2002 ~ dist_lig*treatment, data = main_teso)

summary(m1)
summary(m2)

p <- ggplot(main_teso, aes(x = dist_lig, y = bs_questionnaire_result)) +
  geom_point(aes(color = factor(treatment, levels = c(0,1),
                                labels = c("Control","Treatment"))),
             alpha = 0.6) +
  geom_smooth(aes(color = factor(treatment, levels = c(0,1),
                                 labels = c("Control","Treatment")),
                  linetype = factor(treatment, levels = c(0,1),
                                    labels = c("Control","Treatment"))),
              method = "lm", se = FALSE, linewidth = 1) +
  labs(
    x = "Distance to lightning (km)",
    y = "Collaboration",
    color = "Group",
    linetype = "Group",
    title = "Distance vs. 2002 Test Score by Treatment"
  ) +
  theme_minimal()

p
```

DiD maybe worked?

```{r}
library(readr)

df_long <- main |> filter(teso == 1) |> pivot_longer(cols = c(score_2001, score_2000, score_2002), names_to = "year", values_to = "score")

df_DiD <- df_long |> mutate(year = parse_number(year)) |> mutate(period = as.integer(year >= 2002))

m_did <- lm(score ~ period * treatment * dist_lig, data = df_DiD)
summary(m_did)

d01 <- subset(df_DiD, year %in% c(2000, 2001))
d01$post <- as.integer(d01$year == 2001)
m01 <- lm(score ~ post*treatment*dist_lig , data = d01)
summary(m01)   # coef on post:treatment = ATT for 2001

# 2002 vs 2000
d02 <- subset(df_DiD, year %in% c(2000, 2002))
d02$post <- as.integer(d02$year == 2002)
m02 <- lm(score ~ post*treatment*dist_lig, data = d02)
summary(m02)
```
